---
title: "Problem Set 4"
subtitle: "Behavioral Economics, Boston College"
date: "Nov. 24, Fall 2019"
author: "Sherry Peng TIAN"
output:
  html_document:
  theme: lumen
highlight: pygment
number_sections: TRUE
---
  
<p style="color:red">
*The assignment is worth **100 points**. There are ** 16 questions**. You should have the following packages installed:*
</p>

```{r setup, results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)
library(kableExtra)
library(margins)
```

You may need to update your version of R and follow these steps to download [`margins`](https://github.com/leeper/margins):

```{r,eval=F}
if (!require("remotes")) {
  install.packages("remotes")
  library("remotes")
}
install_github("leeper/prediction")
install_github("leeper/margins")
```

In this problem set you will summarize the paper ["Imperfect Public Monitoring with Costly Punishment: An Experimental Study"](https://www.aeaweb.org/articles?id=10.1257/aer.102.7.3317) (Ambrus and Greiner, AER 2012) and recreate some of its findings. 


# Big picture

**1. What is the main question asked in this paper?**

Even though we already know that costly punishment increases coorperation in finite-horizon social delemma situations, but how the option of costly punishment will affect welfare to people in a more realistic environment, in which people observe each others' behaviors and decisions with some small noise. 

**2. Summarize the experiment design.**

The experiment is a 50-round repeated public good game conducted within a 3X2 factorial design, which includes 
    - three punishment groups: no punishment, regular punishment of 3 tokens each time with a max boundary, and strong punishment of 6 tokens each time without any restrictions 
    - and two noise groups: no noise when disclosing all correct information to subjects and small noise when there is only a public record, a 10% probability of not disclosing the correct "contributed" results. 

**3. Summarize the main results of the experiment.**

**4. Why are these results valuable? What have we learned? Motivate your discussion with a real-world example. In particular discuss the tradeoffs to transparency in groups and how these tradeoffs might be navigated in a firm, or more broadly, a society.**

**5. If punishment is ineffective under imperfect monitoring, what else can you lean on to ensure people cooperate (at least a little) in a public goods problem?**


estimated marginal changes? 
linear -> coeffinicient in the partial derivate is not isolated. 
non-linear regression, prohbitated model; derivie and solve for non-linear 
# Theory

**Payoffs to agent $i$ are**

$$
\pi_i = (e_i - x_i) + \alpha \sum_{i=1}^n x_i
$$
**where $e_i$ is the agent's endowment, $x_i$ is her contribution to the public good, $\alpha$ is the marginal per capita return, and $n$ is the group size.**

**5. Explain $\alpha$ and why in public goods game requires $\frac{1}{n} < \alpha < 1$.**

**6. Suppose $e_i = e = 20$ (i.e. everyone has 20), $\alpha = 0.4$ and $n=4$. Show that $x_i = 0$ is a symmetric Nash equilibrium, but $x_i=20$ is the social optimum. (Recall that in a Nash equilibrium $i$ cannot increase her payoff by changing her contriubtion.)**

# Replication

## Description

<p style="color:red">
*Use `theme_classic()` for all plots.*
<p style="color:red">

**7. Recreate Table 1 and use [`kable()`](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html) to make a publication-quality table (in HTML).**

```{r}
# your code here
```

**8. Recreate Figure 1.**

```{r}
# your code here
```

**9. Recreate Figure 2.**

```{r}
# your code here
```

**10. Recreate Figure 4.**

```{r}
# your code here
```


## Inference

**Consider the linear model**

$$
y = \alpha + \beta_1 x_1 + \beta_2 x_2 + \varepsilon
$$

**11. Write down the marginal effect of $x_1$.**

**Now suppose you have a non-linear model**

$$
y = F(\alpha + \beta_1 x_1 + \beta_2 x_2 + \varepsilon)
$$

**where $F(\cdot)$ is a "link function" that compresses the inputs so that the output $\hat{y} \in [0,1]$.**

**12. Write down the marginal effect of $x_1$. How does this compare to the marginal effect in the linear model?**

**13. A probit model uses the Normal CDF $\Phi$ as the link function, where $\Phi' = \phi$ is the Normal PDF. Use `glm()` to estimate Model 1 in Table 2 (remember to cluster the standard errors at the group level). Assign the model to the object `m1`. Use `summary()` to view the coefficients. Interpret the coefficients. (For more on the probit model, see the appendix.)**

```{r}
# your code here
```


**14. Table 2 reports the average marginal effects (AMEs) of the variables on $P(\text{contribute})$. Calculate the AME to the variable `round` as follows:**

1. **Use [`predict()`](https://www.rdocumentation.org/packages/stats/versions/3.6.1/topics/predict)to create an object `predictions` that contains the predicted z-scores. (i.e. $\hat{\mathbf{X}\beta}$. Hint: use the option `type="link"` in `predict()`.)** 

```{r}
# your code here
```

2. **Use [`dnorm()`](https://www.r-bloggers.com/normal-distribution-functions/) to calculate the probabilities of the predicted z-scores and store the output in an object called `index`.**

```{r}
# your code here
```

3. **Now calculate the marginal effects by multiplying the predicted probabilities times the estimated coefficient for `round` and store the output in `dydxround`.**

```{r}
# your code here
```

4. **Use `mean()` to calculate the AME.**

```{r}
# your code here
```


**15. Verify your calculations with `margins()`, the plot the AMEs. (Note: these will not be exactly the same as those in the paper, since the paper uses an outdated method in Stata.**

```{r}
# your code here
```

**16. Interpret the AMEs.** 

# Appendix: the probit model

Suppose we have latent response variable 

$y^* = \mathbf{X}\beta + \varepsilon$ 

where $\mathbf{X}$ is a $k \times 1$ vector of features $[x_1 \; x_2 \; \dots \; x_k]$ and $\beta$ is a $1 \times k$ coefficient vector. 

The observable binary variable $y$ is defined as 

\begin{align*}
y &= 1 \quad \text{if} \quad  y^* > 0 \\
y &= 0 \quad \text{if} \quad  y^* \leq 0 \\
\end{align*}

If we assume that $\varepsilon \sim N(0,1)$ then 

\begin{align*}
P(y^* > 0)  &= P(\mathbf{X}\beta + \varepsilon > 0)\\
            &= P(\varepsilon > -\mathbf{X}\beta)\\
            &= P(\varepsilon < \mathbf{X}\beta) \quad \text{By symmetry of standard normal}\\
            &= \Phi(\mathbf{X}\beta)
\end{align*}

So $\mathbf{X}\beta$ are z-scores:

\begin{align*}
P(y = 1) &= P(y^* > 0) = \Phi(z \leq \mathbf{X}\beta)\\
P(y = 0) &= P(y^* \leq 0) = 1 - \Phi(z \leq \mathbf{X}\beta)
\end{align*}

where $\Phi$ is the Standard Normal CDF (e.g. $\Phi(0) = 0.5$; half the standard normal distribution lies below $\mu = 0$). 

If we relax the assumption that the error is standard Normal and instead allow it be $\varepsilon \sim N(0, \sigma^2)$, then 

\begin{align*}
P(y^* > 0)  &= P(\mathbf{X}\beta + \varepsilon > 0)\\
            &= P(\frac{\varepsilon}{\sigma} > \frac{-\mathbf{X}\beta}{\sigma})\\
            &= P(\frac{\varepsilon}{\sigma} < \frac{\mathbf{X}\beta}{\sigma}) \\
            &= \Phi(\frac{\mathbf{X}\beta}{\sigma})
\end{align*}

but we cannot estimate $\beta$ and $\sigma$ separately since the probability depends exclusively on their ratio. The standard approach is assume $\sigma = 1$ so $\varepsilon$ is a standard normal. 

