---
title: "R Markdown Notes"
subtitle: ""
date: ""
author: "Lawrence De Geest"
output:
  html_document:
    theme: lumen
    highlight: pygment
    number_sections: TRUE
---

*For a complete introduction see the [official RMarkdown page](https://rmarkdown.rstudio.com/lesson-1.html). Check out the [cheatsheet](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf) as well.*

# Markdown

[Markdown](https://www.markdownguide.org/getting-started/) is a markup language. Formatting is coded and shown when the document is compiled. 

To format inline code, wrap it in two backquotes (e.g. `plot()`). 

To display code blocks, indent twice:

    plot()

Quote with `>`:

>It was the best of times, it was the worst of times.

Wrap a word in one asterisks (`*`) to make it *italic*, two asterisks (`**`) to make it **bold**, three to make it ***bold italics***. 

Unordered lists:

* Item one
* Item two 
* Item three

Ordered lists:

1. Item one
2. Item two
3. Item three

Hashtags (`#`) are used for headers, `##` for subheaders, `###` for subheaders, and so on.

## Subheader

## Subsubheader

# Code chunks

R code can be placed in "chunks":

```
{r}
```

Inside `{r}` you can add options like:

* `eval=F` so the code is printed but not evaluated;
* `echo=F` so the code is evaluated but not printed;
* a name for the chunk (e.g. `{r demo}`)
* and [more](https://yihui.name/knitr/options/?version=1.1.383&mode=desktop).



## Figures

```
{r, barplot, fig.align="center", fig.width=6, fig.height=6, fig.cap="Figure caption."}
```

```{r, barplot, message=FALSE, warning=FALSE,fig.align="center", fig.width=6, fig.height=6, fig.cap="Figure caption."}
library(tidyverse)
mpg %>%
  ggplot(aes(reorder(class, hwy), hwy)) + 
  stat_summary(fun.y = mean, geom = "bar", fill="tomato2") + 
  stat_summary(fun.data = mean_se, geom = "errorbar", width=0.1) + 
  labs(y="Miles per gallon (highway)", x="Car class", title="Title", subtitle = "Subtitle") + 
  theme_classic()
```

## Tables

There are lots of packages to [format tables](https://rmarkdown.rstudio.com/lesson-7.html). Personally I like [`library(kableExtra)`](http://haozhu233.github.io/kableExtra/awesome_table_in_html.html).

```{r table, warning=F}
library(kableExtra)
mpg %>% 
  group_by(class) %>% 
  summarise(avg_hwy = mean(hwy), avg_city = mean(cty)) %>% 
  kable() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive", full_width = F))
```

Use [`library(stargazer)`](https://cran.r-project.org/web/packages/stargazer/index.html) for regression tables. All the options are in this [document](https://cran.r-project.org/web/packages/stargazer/stargazer.pdf). Here is a nice [cheatsheet](https://www.jakeruss.com/cheatsheets/stargazer/).


```{r regression_table, warning=F, message=F, results="asis"}
library(stargazer)
m1 <- lm(hwy~class, data=mpg)
m2 <- lm(cty~class, data=mpg)
stargazer(m1,m2, 
          type="html", 
          style = "aer",
          omit.stat = c("adj.rsq", "f", "ser"),
          dep.var.labels.include = FALSE,
          column.labels = c("MPG (Highway)", "MPG (City)"),
          covariate.labels = c("Compact", "Midsize", "Minivan",
                               "Pickup", "Sub-compact", "SUV"),
          title = "Regression results",
          notes = "*Source*: Own calculations."
          )
```

# Math

Math is written with [LaTeX](https://en.wikibooks.org/wiki/LaTeX/Mathematics). Just like with HTML (e.g. `<p> ... </p>`) you wrap math code in "environments" (e.g. `\begin{equation} ... \end{equation}`.)

To write an inline equation, wrap it in `$`. `$y = \beta X$` produces $y=\beta X$. 

To write an full-length equation, wrap it inside `\begin{equation} ... \end{equation}`:

    \begin{equation}
    y=\beta X + \varepsilon
    \end{equation}

\begin{equation}
y=\beta X + \varepsilon
\end{equation}

To align equations use `aligned`:

    \begin{aligned}
    y &=\beta X + \varepsilon \\
      &= \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \varepsilon
    \end{aligned}

\begin{aligned}
y &=\beta X + \varepsilon \\
  &= \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \varepsilon
\end{aligned}

Use `\frac{}{}` to write fractions: `\frac{\partial f}{\partial x}` produces $\frac{\partial f}{\partial x}$